Mixed effect tweedie model in TMB

### 2019-04-15 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
testing batch processes for sdmTMB_cv

# Fit spatial temporal models to bottom temperature

```{r setup}
library(tidyverse)
library(sdmTMB)
library(future)

d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
#set.seed(39201114) # original seed, predictions max out at .88 cor with true temp
set.seed(392011149)
#d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
d_trawl$depth_scaled3 <- (d_trawl$depth_scaled)^3
```

### Functions for model exploration:
```{r}
generate_cv_args <- function(
                formula = formula,
                time_varying = time_varying,
                ar1_fields = c(FALSE,TRUE), 
                include_spatial = c(FALSE,TRUE),
                n_knots = c(30,80) ){
  arguments <- expand.grid(
    formula = formula,
    time_varying = time_varying,
    include_spatial = include_spatial,
    ar1_fields = ar1_fields,
    n_knots = n_knots
  )
  args <- as_tibble(arguments)
  args
}

plot_cv_meshes <- function(cv_out) {
  k <- max(cv_out$data$cv_fold)
  op <- graphics::par(
      mfrow = c((sqrt(k)), ceiling(sqrt(k))),
      mar = c(1, 1, 1, 1)
    )
  for (i in seq_len(k)){
    plot_spde(cv_out$models[[i]]$spde)
  }
graphics::par(op)
}

aic <- function(m) {
  k <- length(m$model$par)
  nll <- m$model$objective
  2 * k - 2 * (-nll)
}
```

# 10-fold cross-validation 
#FIXME: need to extract information on whether or not models converge
## All fixed slope models
```{r}
cv_args <- generate_cv_args (formula = list(
    temperature_c ~ 0 + as.factor(year) + depth_scaled,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3),
    time_varying = list(NULL),
    ar1_fields = c(FALSE, TRUE), 
    include_spatial = c(FALSE,TRUE),
    n_knots = c(30,80,150)
  )

plan(multiprocess)
sdmTMB_cv_batch1 <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl)
cv_args1 <- cv_args
cv_loglik <- c()
for (i in seq_len(nrow(cv_args1))) {
  cv_loglik[[i]] <- round(sdmTMB_cv_batch1[[i]]$sum_loglik, 1)
}
cv_args1$formula <- as.character(cv_args1$formula)
cv_args1$cv_loglik <- cv_loglik
cv_args_notv <- as.data.frame(cv_args1) %>% select(., -time_varying)
# View(cv_args_notv)
# write.csv(cv_args_notv, file = "cv-all-non-time-varying.csv")
# saveRDS(sdmTMB_cv_batch1, file = "cv-all-non-time-varying.rds")
glimpse(cv_args_notv)
```

## All time-varying slope models
```{r}
cv_args <- generate_cv_args (formula = list(
    temperature_c ~ 0 + as.factor(year)),
    time_varying = list(
    ~ 0 + depth_scaled,
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
    ar1_fields = c(FALSE, TRUE), 
    include_spatial = c(FALSE,TRUE),
    n_knots = c(30,80,150)
    )

plan(multiprocess)
sdmTMB_cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl)
cv_args2 <- cv_args
cv_args2$cv_loglik <- c()
for (i in seq_len(nrow(cv_args))) {
  cv_args2$cv_loglik[[i]] <- c(sdmTMB_cv_batch[[i]]$sum_loglik)
}
cv_args2$formula <- as.character(cv_args$formula)
cv_args2$time_varying <- as.character(cv_args$time_varying)
cv_args_tv <- as.data.frame(cv_args2) 
# View(cv_args_tv)
# write.csv(cv_args_tv, file = "cv-all-time-varying.csv")
# saveRDS(sdmTMB_cv_batch, file = "cv-all-time-varying.rds")
glimpse(cv_args_tv)
```

## vary n_knots for top model
```{r}
cv_args_knots <- generate_cv_args (formula = list(
    temperature_c ~ 0 + as.factor(year)),
    time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
    ar1_fields = c(FALSE,TRUE), 
    include_spatial = c(FALSE,TRUE),
    n_knots = c(60,90,110,130,160)
    )

plan(multiprocess)
sdmTMB_cv_knots2 <- purrr::pmap(cv_args_knots, sdmTMB_cv, data = d_trawl)
cv_args3 <- cv_args_knots
cv_args3$cv_loglik <- c()
for (i in seq_len(nrow(cv_args3))) {
  cv_args3$cv_loglik[[i]] <- c(sdmTMB_cv_knots2[[i]]$sum_loglik)
}
cv_args3$formula <- as.character(cv_args3$formula)
cv_args3$time_varying <- as.character(cv_args3$time_varying)
cv_args_knots2 <- as.data.frame(cv_args3) 
# View(cv_args_knots2)
# write.csv(cv_args_knots2, file = "cv-varying_knots2.csv")
# saveRDS(sdmTMB_cv_knots2, file = "cv-varying_knots2.rds")
glimpse(cv_args_knots2)
```

Load saved batches:
```{r}
non_time_varying_batch <- readRDS(file = "cv-all-non-time-varying.rds")
time_varying_batch <- readRDS(file = "cv-all-time-varying.rds")
knot_varying_batch <- readRDS(file = "cv-varying_knots.rds")
knot_varying_batch2 <- readRDS(file = "cv-varying_knots2.rds")
```

Current top model
```{r}
plan(multiprocess)
temp3_rw <- sdmTMB_cv(data = d_trawl, n_knots = 130, x = "X", y ="Y", k_folds = 10, 
                   formula = temperature_c ~ 0 + as.factor(year),
                   time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
                   time = "year", family = gaussian(link = "identity"), 
                   ar1_fields = TRUE, include_spatial = TRUE)
temp3_rw$models
sum(temp3_rw$sum_loglik)
```

>>>>

---
output: html_document
editor_options: 
  chunk_output_type: inline
---
# Fit spatial temporal models to bottom temperature

```{r setup}
library(tidyverse)
library(sdmTMB)
library(future)
plan(multiprocess)
d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
#d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
d_trawl$depth_scaled3 <- (d_trawl$depth_scaled)^3
unique(d_trawl$year)
```

### Functions for model exploration:
```{r}
aic <- function(m) {
  k <- length(m$model$par)
  nll <- m$model$objective
  2 * k - 2 * (-nll)
}

generate_cv_args <- function(
                formula = formula,
                time_varying = time_varying,
                ar1_fields = c(FALSE), # simiple, fast defaults
                include_spatial = c(FALSE), # simiple, fast defaults
                n_knots = c(100), # simiple, fast defaults
                seed = c(999)
  ){
  arguments <- expand.grid(
    formula = formula,
    time_varying = time_varying,
    include_spatial = include_spatial,
    ar1_fields = ar1_fields,
    n_knots = n_knots,
    seed = seed
  )
  #browser()
  as.data.frame(arguments)
}

sdmTMB_cv_batch <- function (cv_args, data, ...) {
  cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl, ...)
  cv_out <- cv_args
  cv_out$cv_loglik <- c()
  cv_out$all_converged <- c()
  cv_out$max_grad <- c()
  for (i in seq_len(nrow(cv_args))) {
    cv_out$cv_loglik[[i]] <- round(cv_batch[[i]]$sum_loglik, 1)
    cv_out$all_converged[[i]] <- all(cv_batch[[i]]$converg)
    cv_out$max_grad[[i]] <- max(cv_batch[[i]]$max_grad)
    if (is.null(cv_out$time_varying[[i]])) 
      cv_out$time_varying[[i]] <- as.character(NA)
  }
  cv_out$formula <- as.character(cv_out$formula)
  cv_out$time_varying <- as.character(cv_out$time_varying)
  list(summary = as.data.frame(cv_out), cv_model_set = cv_batch)
}

plot_cv_meshes <- function(sdmTMB_cv_output) {
  cv_output <- sdmTMB_cv_output
  k <- max(cv_output[["data"]]$cv_fold)
  if (k == 2 | k == 5 | k == 10) {
    op <- graphics::par(
      mfrow = c((sqrt(k)), ceiling(sqrt(k))),
      mar = c(1, 1, 1, 1)
      )
  } else { 
    op <- graphics::par(
      mfrow = c(ceiling(sqrt(k)), ceiling(sqrt(k))),
      mar = c(1, 1, 1, 1)
    )
  }
  for (i in seq_len(k)){
    plot_spde(cv_output$models[[i]]$spde)
  }
  graphics::par(op)
}

```

Short example: 
```{r, eval=FALSE}
cv_args <- generate_cv_args (formula = list(
    temperature_c ~ 0 + as.factor(year) + depth_scaled),
    time_varying = list(NULL),
    ar1_fields = c(FALSE), 
    include_spatial = c(FALSE),
    n_knots = c(30), 
    seed = c(111,222)
  )

cv <- sdmTMB_cv_batch(cv_args, d_trawl, k_folds = 4, time = "year", x = "X", y = "Y")
glimpse(cv$summary)
plot_cv_meshes(cv$cv_model_set[[1]])
plot_cv_meshes(cv$cv_model_set[[2]])
```

# 10-fold cross-validation

## Just compare fixed slope models with simplest random effect structure
```{r}
cv_args <- generate_cv_args (formula = list(
    temperature_c ~ 0 + as.factor(year) + depth_scaled,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3),
    time_varying = list(NULL),
    ar1_fields = c(FALSE), 
    include_spatial = c(FALSE),
    n_knots = c(150), #start with close to projected optimum n_knots 
    seed = c(1111,2221,3331)
  )

non_time_varying <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
View(non_time_varying$summary)
write.csv(non_time_varying$summary, file = "cv-non-time-varying.csv")
saveRDS(non_time_varying, file = "cv-non-time-varying.rds")
glimpse(non_time_varying$summary)
```

## Compare time-varying slope models still with simplest random effect structure
```{r}
cv_args <- generate_cv_args (formula = list(
    temperature_c ~ 0 + as.factor(year)),
    time_varying = list(
    ~ 0 + depth_scaled,
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
    ar1_fields = c(FALSE), 
    include_spatial = c(FALSE),
    n_knots = c(150), # keep same n_knots
    seed = c(1111,2221,3331) # keep same seeds
    )

time_varying  <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")

View(time_varying$summary)
write.csv(time_varying$summary, file = "cv-time-varying.csv")
saveRDS(time_varying, file = "cv-time-varying.rds")
glimpse(time_varying$summary)
```

## Now vary n_knots and random effect configuation for top model(s) 
```{r}
cv_args <- generate_cv_args(formula = list(
    temperature_c ~ 0 + as.factor(year)),
    time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
    ar1_fields = c(FALSE, TRUE), 
    include_spatial = c(FALSE, TRUE),
    n_knots = c(85,110,135),
    seed = c(1111,2221,3331)
    )

vary_rf_knots <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")

View(vary_rf_knots$summary)
write.csv(vary_rf_knots$summary, file = "cv-varying-rf.csv")
saveRDS(vary_rf_knots, file = "cv-varying-rf.rds")
glimpse(vary_rf_knots$summary)
```

## Change seeds 
```{r}
cv_args <- generate_cv_args(formula = list(
    temperature_c ~ 0 + as.factor(year)),
    time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
    ar1_fields = c(FALSE), 
    include_spatial = c(TRUE),
    n_knots = c(120),
    seed = c(4441,5551,6661,7771,8881,9991)
    )

more_seeds <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")

View(more_seeds$summary)
write.csv(more_seeds$summary, file = "cv-more-seeds.csv")
saveRDS(more_seeds, file = "cv-more-seeds.rds")
glimpse(more_seeds$summary)
```

Load saved batches:
```{r}
# non_time_varying <- readRDS(file = "cv-non-time-varying.rds")
# time_varying <- readRDS(file = "cv-time-varying.rds")
# vary_rf <- readRDS(file = "cv-varying-rf.rds")
# knot_refining <- readRDS(file = "cv-refine-knots.rds")

# non_time_varying <- read.csv(file = "cv-non-time-varying.csv")
# time_varying <- read.csv(file = "cv-time-varying.csv")
# vary_rf <- read.csv(file = "cv-varying-rf.csv")
# knot_refining <- read.csv(file = "cv-refine-knots.csv")

all_cv_loglik <- rbind(non_time_varying$summary, time_varying$summary, non_tv$summary, tvar$summary, vary_rf$summary, vary_rf_knots$summary, knot_refining$summary, more_seeds$summary, deparse.level = 1, make.row.names=TRUE)
View(all_cv_loglik)

by_seed <- all_cv_loglik %>% group_by(seed) %>% 
  mutate(., max_ll_per_seed = max(cv_loglik)) %>% 
  mutate(., delta_ll = cv_loglik - max_ll_per_seed)

View(by_seed)
all_cv_loglik <- by_seed %>% ungroup()

write.csv(all_cv_loglik, file = "all_cv_loglik.csv")
```

Current top model
```{r}
plan(multiprocess)
temp3_rw <- sdmTMB_cv(data = d_trawl, n_knots = 130, x = "X", y ="Y", k_folds = 10, 
                   formula = temperature_c ~ 0 + as.factor(year),
                   time_varying = ~ 0 + depth_scaled + depth_scaled2,
                   time = "year", family = gaussian(link = "identity"), 
                   ar1_fields = FALS, include_spatial = TRUE)
temp3_rw$models
sum(temp3_rw$sum_loglik)
```
